MapReduce to My Ears\\ {\Large Using Big Data for Music Synthesis}

### 2013-05-10 writeup/paper.tex
adding stuff to the conclusion

\begin{figure}
\begin{lstlisting}[language=Java]
public class WriteImagesToSequenceFile
{
  BufferedImage img = null;
  SequenceFile.Writer writer = null;
  FileSystem fs = 
      FileSystem.get(conf);
  FileStatus[] fss = 
    fs.globStatus(new Path(input 
      + "/*.jpg"));
  writer = SequenceFile
      .createWriter(fs, 
    conf,outPath,IntWritable.class,
    ArrayListOfIntsWritable.class);
  int counter = 0;
  for (FileStatus status : fss) {
    IntWritable image_counter = 
      new IntWritable();
    ArrayListOfIntsWritable 
      pixelData = 
        new ArrayListOfIntsWritable();
    img = ImageIO.read(
      new File(status.getPath()
           .toString().substring(5)));
    // Get RBG integer for each pixel 
    for(int j = 0; j < thumbs; j++) {
      for(int i = 0; i< thumbs; i++) {
        pixelData
          .add(img.getRGB(i,j));
      }
    }
    // Write to the sequence file
    image_counter.set(counter);
    writer.append(image_counter,
      pixelData);
    counter += 1;
  } 
}
\end{lstlisting}
\caption{Snippet of code from \code{WriteToSequenceFile} showing
  format used in sequence files throughout the MapReduce jobs.}
\end{figure}

@@

\begin{figure}
\label{figure:colortomusic-mapper}
\begin{lstlisting}[language=Java]
public void map(IntWritable key, 
    ArrayListOfIntsWritable pixels,
    Context context)
  throws IOException, InterruptedException {
  int regionCounter = 1;
  MidiNote midiNote = null;
  for(int i = 0; i < pixels.size(); i++){
    pixel = pixels.get(i);
    // Conversion from color to note
    midiNote = Color2Music
      .convert(new Color(pixel));
    REGION_IMAGENO.set(regionCounter,
      key.get());
    NOTE_VELOCITY.set(midiNote.getTone(),
      midiNote.getVelocity());
    context.write(REGION_IMAGENO, NOTE_VELOCITY);
    regionCounter++;
  }
}
\end{lstlisting}
\caption{A portion of our mapper that takes color data to MIDI notes.
  For each pixel in the original image, the conversion in
  \code{Color2Music} is applied before setting the output key (as a
  region, number pair).}
\end{figure}

@@

The result of our mapreduce pipeline is a set of MIDI music: one for
each region in the image set.  To turn these into music that can be
played, we use mixing software (such as GarageBand) to mix the regions
into full scores.

@@

\subsection{Sample datasets}

To test the system, we use a variety of different sets of images.
These images were obtained from
Flickr\footnote{\url{http://www.flickr.com/}}, using a public API and
a Python script \cite{flickerpy} to get test data.

>>>>

\begin{figure}
\begin{lstlisting}[language=Java]
public class WriteImgsToSequence
{
  BufferedImage img = null;
  SequenceFile.Writer writer=null;
  FileSystem fs = 
      FileSystem.get(conf);
  FileStatus[] fss = 
    fs.globStatus(new Path(input 
      + "/*.jpg"));
  writer = SequenceFile
      .createWriter(fs, 
    conf,outPath,IntWritable.class,
    ArrayListOfIntsWritable.class);
  int counter = 0;
  for (FileStatus status : fss) {
    IntWritable image_counter = 
      new IntWritable();
    ArrayListOfIntsWritable 
     pixelData = 
      new ArrayListOfIntsWritable();
    img = ImageIO.read(
      new File(status.getPath()
         .toString().substring(5)));
    // Get RBG for each pixel 
    for(int j = 0; j<thumbs; j++) {
     for(int i = 0; i<thumbs; i++) {
        pixelData
          .add(img.getRGB(i,j));
      }
    }
    // Write to the sequence file
    image_counter.set(counter);
    writer.append(image_counter,
      pixelData);
    counter += 1;
  } 
}
\end{lstlisting}
\caption{Snippet of code from \code{WriteToSequenceFile} showing
  format used in sequence files throughout the MapReduce jobs.}
\end{figure}

@@

\begin{figure}
\label{figure:colortomusic-mapper}
\begin{lstlisting}[language=Java]
public void map(IntWritable key, 
    ArrayListOfIntsWritable pixels,
    Context context)
{
  int regionCounter = 1;
  MidiNote midiNote = null;
  for(int i=0; i<pixels.size();i++){
    pixel = pixels.get(i);
    // Conversion from color to note
    midiNote = Color2Music
      .convert(new Color(pixel));
    REGION_IMAGENO.set(regionCounter,
      key.get());
    NOTE_VELOCITY.set(midiNote.getTone(),
      midiNote.getVelocity());
    context.write(REGION_IMAGENO,
       NOTE_VELOCITY);
    regionCounter++;
  }
}
\end{lstlisting}
\caption{A portion of our mapper that takes color data to MIDI notes.
  For each pixel in the original image, the conversion in
  \code{Color2Music} is applied before setting the output key (as a
  region, number pair).}
\end{figure}

@@

The result of our mapreduce pipeline is a set of notes in MIDI: one
for each region in the image set.  We postprocess the MapReduce output
by taking it off HDFS using a script and converting to the MIDI format
(adding on a header about the number of parts and other small details
to be a valid MIDI file). To turn these into music that can be played,
we use mixing software (such as GarageBand) to mix the regions into
full scores and generate MP3 files.

To visualize music and colors, we take the resulting music and mix it
with a video generated from the thumbnailed images in our input
dataset.  We do this using a process similar to preprocessing: we
simply blow up the thumbnailed images to a larger size, concatenate
them together into subsequent frames of a video, and mix the video
with the music generated from our MapReduce phases.

@@

To test the system, we use a variety of different sets of images.
These images were obtained from
Flickr\footnote{\url{http://www.flickr.com/}} and \code{memebase.com},
using a public API and a Python script \cite{flickerpy} to get test
data.  We converted animated GIFs into individual frames again using a
preprocessing phase.  Our MapReduce code takes as input simply a
\code{SequenceFile} (as discussed in Section~\ref{sec:conversion}), so
we simply feed in a folder of JPG images to our preprocessing phase.

As a result, we produced various pieces of music.  Our longest piece
(named \code{Digital Nightmare}) was over six minutes long and
composed of over 3000 images.  We also produced shorter (self
contained) pieces from various related images: such as GIFs of cat
jumping animations.  All of these pieces are made available in our
repository.

Our MapReduce phases ran very quickly (on the order of minutes, mostly
startup time).  We believe that the reason for this is that by the
time preprocessing was completed we ended up with a relatively small
amount of data to be fed into MapReduce.  Were we to do the project
over again we would have fed in much more data to the MapReduce jobs
and optimized the selection of regions (from each image) to be the
``best sounding'' piece of music that could be produced.  We thought
about how to devise a measure of ``realistic'' music, but could not
easily see one.  One potential next application would be to use Hadoop
to mine this correlation.  The idea is that realistic music generally
has some recurring nature (backbeat, for example) even if very
complex.  Therefore, we would want to apply a penalty term to our
measure whenever we generated music that had too complex a principal
structure (this would avoid creating chaotic sounding music).  We did
some initial work on this (writing down some equations, studying how
to analyze MIDI for it's beat) but didn't have anything concrete by
the time we needed to do implementation.

We also noticed that there might be some interesting problems in
taking a song and a set of videos and finding out which best matched
each other.  (In essence, this would let you search for image based on
video.)  We started to write an inverse transformation within
MapReduce\footnote{src/main/phase2/Color2Music} to explore this idea,
but haven't finished doing the work yet.
